---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(tidyverse)
library(ppendemic)
```

# ppendemic 


<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![CRAN status](https://www.r-pkg.org/badges/version/ppendemic)](https://CRAN.R-project.org/package=ppendemic)
[![](http://cranlogs.r-pkg.org/badges/grand-total/ppendemic?color=green)](https://cran.r-project.org/package=ppendemic)
[![](http://cranlogs.r-pkg.org/badges/last-week/ppendemic?color=green)](https://cran.r-project.org/package=ppendemic)

[![R-CMD-check](https://github.com/PaulESantos/ppendemic/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/PaulESantos/ppendemic/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/PaulESantos/ppendemic/branch/main/graph/badge.svg)](https://app.codecov.io/gh/PaulESantos/ppendemic?branch=main)
<!-- badges: end -->

## Overview

`ppendemic` is an R package that provides easy access to a new database containing the list of endemic plant species found in Peru. The package includes a collection of 7,249 species with detailed botanical information, such as accepted names, family, authorship, publication details, and the place and date of first publication.

## Installation

You can install the `ppendemic` package from GitHub using the following command:

```r
pak::pak("PaulESantos/ppendemic")
```


## Data base 

The ppendemic database consists of 7,249 rows and 7 columns, with each row representing a unique endemic plant species. The columns in the database include:

1. `accepted_name`: The accepted scientific name of the endemic plant species.
2. `accepted_family`: The family to which the plant species belongs.
3. `accepted_name_author`: The author of the accepted scientific name.
4. `publication_author`: The author of the publication where the species was described.
5. `place_of_publication`: The place of publication of the species.
6. `volume_and_page`: The volume and page number of the publication.
7. `first_published`: The year of the first publication.

The database covers a total of 169 families, and it is particularly rich in Orchidaceae, Asteraceae, Piperaceae, Solanaceae, Fabaceae, Bromeliaceae, Melastomataceae, Araceae, Rubiaceae, and Gentianaceae families, which have the highest number of endemic species in Peru.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
ppendemic::ppendemic_tab |> 
  group_by(accepted_family) |> 
  summarise(n_especies = n_distinct(accepted_name)) |> 
  arrange(desc(n_especies)) |> 
  top_n(10) |> 
  ggplot(aes(fct_reorder(accepted_family, n_especies, .desc = TRUE),
             n_especies)) +
  geom_col() +
  labs(x = "Families",
       y = "Number of species") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  theme_bw()
```


## Citation

To cite the `ppendemic` package, please use:

```{r}
citation("ppendemic")
```

