---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ppendemic

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/ppendemic)](https://CRAN.R-project.org/package=ppendemic)
<!-- badges: end -->

El objetivo de ppendemic es brindar acceso a la información de "El libro rojo de las plantas endémicas del Perú. Blanca León et al 2006", mediante un grupo de herramientas que permiten evaluar si nuestros registros contienen especies endémicas del Perú. 

The goal of ppendemic is to gives access to the "El libro rojo de las plantas endémicas del Perú. Blanca León et al 2006", with a set of functions to validate if you have endemic plants in your plants records.

## Installation

You can install the released version of ppendemic from [github](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("PaulESantos/ppendemic")
```

## Ejemplo:
Pep_check, es la función básica de ppendemic:


```{r example}
library(tidyverse)
library(ppendemic)
## basic example code
```


```{r spp}
spp <- c("Mauria denticulata", "Mauria flexuosa")

pep_check(spp)

spp1 <- enframe(spp)
spp1 %>% 
  mutate(endemic = pep_check(value))
```
